---
title: "Homework 10"
author: "Kimberly McNeil"
date: "4/15/2026"
output:
  html_document:
    code_folding: show
linkcolor: blue
urlcolor: blue
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(
  echo = TRUE#, 
  # warning=FALSE, 
  # message=FALSE
  )
library(knitr)
library(tidyverse)
library(dplyr)
library(lubridate)
```


# General Social Survey

## Recoding Marital Levels

The following table showcases the distribution of respondents across the newly defined marital categories:^[Question 1 Complete]

```{r}
gss_marital = gss_cat %>%
  mutate(
    marital_new = fct_collapse(
      marital,
      "Pre-Marriage" = "Never married",
      "Marriage" = c("Married", "Separated"), 
      "Post-Marriage" = c("Divorced", "Widowed"),
      "Missing" = "No answer"
    )
  )

gss_marital %>%
  count(marital_new) %>%
  kable(caption = "Frequency Table for new Marital Levels")
```

## 

2. Count each year-marriage-status combination. Then take that summary, grouping by year, make a new
column for the proportion of responses in each year that fall into each marriage level. Then store this
data view in an object.

## 

3. Reorder the levels of marriage status by the latest year’s proportion and re-store the object.

##

4. Make a line plot of proportion versus year, colored by your four marriage levels.


# Los Angeles Lakers
We will works with some dates and times from the lubridate::lakers dataset.

## 

1. Make a new column that directly converts the date (integer data) column to a new column of date
data. What goes wrong? Provide a table of the first 6 rows (you may select a subset of the other
columns) and a guess for why the conversion failed.

##

2. Make a new column that converts the date column to strings and then converts the string to date data.
Provide a table of the first 6 rows (you may select a subset of the other columns) and a guess for why
this works better.

##

3. Find the distinct dates in the dataset (Tip: Use the date column you made along with distinct().),
make a new column extracting the day of the week (abbreviated name), and make a bar plot of the
games on different days.
