---
title: "Homework 10"
author: "Kimberly McNeil"
date: "4/15/2026"
output:
  html_document:
    code_folding: show
linkcolor: blue
urlcolor: blue
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(
  echo = TRUE#, 
  # warning=FALSE, 
  # message=FALSE
  )
library(knitr)
library(tidyverse)
library(dplyr)
library(lubridate)
```


# General Social Survey

## Recoding Marital Status

The following table showcases the distribution of respondents across the newly defined marital categories:^[Question 1 Complete]

```{r}
gss_marital = gss_cat %>%
  mutate(
    marital_new = fct_collapse(
      marital,
      "Pre-Marriage" = "Never married",
      "Marriage" = c("Married", "Separated"), 
      "Post-Marriage" = c("Divorced", "Widowed"),
      "Missing" = "No answer"
    )
  )

gss_marital %>%
  count(marital_new) %>%
  kable(caption = "Frequency Table for new Marital Levels")
```

## Yearly Proportions of Marital Status

The following table showcases the first six rows of each year-marriage-status combination's count and calculated proportion:^[Question 2 Complete] 

```{r}
marriage_year_summary = gss_marital %>%
  count(year, marital_new) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n))

kable(head(marriage_year_summary), caption = "First Six Rows of Each Year-Marriage-Status Combination")
```

## Reorder Marital Levels by Latest Year

Now, let's reorder the levels of marriage status by the latest year’s proportion and re-store the object:^[Question 3 Complete] 

```{r}
latest_year = max(marriage_year_summary$year) # select single col --> df$column_name

marriage_year_summary = marriage_year_summary %>%
  mutate(
    marital_new = fct_reorder(marital_new, year)
  )
```


## Line Plot of Proportion Versus Year

The following line plot showcases the marital status' proportions over time in years, colored by our four marriage levels:^[Question 4 Complete] 

```{r}
ggplot(marriage_year_summary, aes(x = year, y = prop, color = marital_new)) +
  geom_line() +
  labs(
    title = "Marital Status Proportions Over Time",
    x = "Year",
    y = "Proportion",
    color = "Marital Status"
  )
```



# Los Angeles Lakers
We will works with some dates and times from the lubridate::lakers dataset.

## 

1. Make a new column that directly converts the date (integer data) column to a new column of date
data. What goes wrong? Provide a table of the first 6 rows (you may select a subset of the other
columns) and a guess for why the conversion failed.

##

2. Make a new column that converts the date column to strings and then converts the string to date data.
Provide a table of the first 6 rows (you may select a subset of the other columns) and a guess for why
this works better.

##

3. Find the distinct dates in the dataset (Tip: Use the date column you made along with distinct().),
make a new column extracting the day of the week (abbreviated name), and make a bar plot of the
games on different days.
